### 2011-06-07 thesis/novelty-intro.tex
added reference to novelty detection on documents.

The ability to identify novel cases is crucial in any autonomous system
that is deployed to unknown or to uncontrolled environments, as it gives the
system the ability to detect something is not conforming to it knowledge and
therefore should proceed with caution.
It has several applications such as fault detection~\cite{tarassenko1999novelty},
intrusion detection~\cite{fan2001using},
detection of masses in mammograms~\cite{tarassenko1995novelty}.
Any normal classification application is also a good candidate for extending
with novelty detection, as it allows to detect samples the system has not been
trained on and therefore the classification is considered
instable~\cite{devarakota2008reliability}.
E.g.\ applied to digit-recognition~\cite{tax1998outlier}
or to detection of novel inputs on a neural network~\cite{bishop1994novelty}.

@@

\section{Novelty Detection as a Threshold}
% Explain threshold approach to novelty detection
Due to the desire of robustness, a novelty detection system is often given the option
to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
between error and detection: the more the system tries to increase detection, the more
prone to produce errors it becomes.

@@

\clearpage
\section{A Practical Example}
\label{sec:unlabelled-data}
In order to sum up the presented concepts on novelty detection with a threshold
function and exemplify how to use graphical models in the context of
multi-modality room classification a synthetic dataset was generated.
The dataset was kept simple by only modelling directly sensed features from a
room, skipping any structural knowledge or sensing model such as room
connectivity and extra hidden variables.

>>>>

The ability to identify novel cases is crucial in any autonomous system
that is deployed to unknown or to uncontrolled environments, as it gives the
system the ability to detect something is not conforming to it knowledge and
therefore should proceed with caution.
It has several applications such as fault detection~\cite{tarassenko1999novelty},
intrusion detection~\cite{fan2001using},
detection of masses in mammograms~\cite{tarassenko1995novelty} or detection of
novel and useful documents~\cite{zhang2002novelty}.

Any normal classification application is also a good candidate for extending
with novelty detection, as it allows to detect samples the system has not been
trained on and therefore the classification is considered
instable~\cite{devarakota2008reliability}.
E.g.\ applied to digit-recognition~\cite{tax1998outlier}
or to detection of novel inputs on a neural network~\cite{bishop1994novelty}.

@@

% TODO: wuld be nice to have this one starting on a new page.. but still make
% sure it looks good and not a forced break.
% Reason to have this on a diferent page is to make it really clear novelty
% content (and not only chit-chat) starts now.
\clearpage
\section{Novelty Detection as a Threshold}
% Explain threshold approach to novelty detection
Due to the desire of robustness, a novelty detection system is often given the option
to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
between error and detection: the more the system tries to increase detection, the more
prone to produce errors it becomes.

@@

\section{A Practical Example}
\label{sec:unlabelled-data}
In order to sum up the presented concepts on novelty detection with a threshold
function and exemplify how to use graphical models in the context of
multi-modality room classification a synthetic dataset was generated.
The dataset was kept simple by only modelling directly sensed features from a
room, skipping any structural knowledge or sensing model such as room
connectivity and extra hidden variables.
